\documentclass[preprint,10pt,nonatbib]{sigplanconf}
\title{Code Spartanization}
\subtitle{\scriptsize \emph{one rational approach for⏎ resolving religious style wars}}
\authorinfo
  {Yossi Gil \and Matteo Orrù}
  {%
    Dept.\@ of Comp.\@ Sc.\@⏎
    The Technion---I.I.T.⏎
    Haifa 32000, Israel
  }
  {❴ YoGi~$|$~Matteo.Orru ❵ @cs.technion.ac.il}

\usepackage{\jobname}
\begin{document}

\maketitle

\def\ignore#1{}
\def\yossi{\marginpar[Y$⇒$]{$⇐$Y}}
\def\matteo{\marginpar[M$⇒$]{$⇐$M}}

\begin{abstract}
  \input{abstract}
\end{abstract}

\section{Introduction}
\input{aa}

\paragraph{Outline.} The present manuscript is organized as it follows. 
\Cref{section:process} describes the code spartanization
process including what we call ‟cleanup”, ‟structural”, and ‟nominal”
spartanization, taking some space to elaborate on rationale and the automation
of the process. \cref{section:empirical} reports the empirical results.
\cref{section:zz} concludes, also with directions for future work. 
% The appendix demonstrates how spartanization interacts with method extraction
% refactoring process.

\section{The Process of Code Spartanization}
\label{section:process}
\input{process}

\section{Empirical Results}
\label{section:empirical}
\input{empirical}

\section{Conclusion and Future Research}
\label{section:zz}
\input{zz}

\paragraph{Acknowledgment.} We thank G. Lalouche for inspiring discussions.

\balance
\bibliographystyle{abbrv}\small
\bibliography{%
other-shorthands,%
author-names,%
publishers-abbreviated,%
conferences-abbreviated,%
journals-abbreviated,%for
journals-full,%
yogi-tr,%
yogi-book,%
yogi-practice,%
yogi-misc,%
yogi-journal,%
yogi-theory,%
yogi-confs,%
\jobname,%
GPCE,%
OOPSLA,%
PLDI,%
USENIX,%
ECOOP%
}
\end{document}

\eject
\appendix
\section{Spartanization and Method Extraction}
\label{section:techniques}
\input{extraction}

